Effect of Black Hole Spin on Christodoulou Memory

### 2019-02-11 notes/Paper.tex
updated paper

Gravitational wave memory effect results in a permanent displacement of test masses after the gravitational wave has passed through an "ideal" GW detector. There are two kinds of gravitational wave memory effects: linear and nonlinear. The linear memory is produced by gravitational sources that produce a net change in the time derivative of one or more of their source-multipole moments. At leading order in a Post-Newtonian (PN) expansion, the linear memory causes a net change in the GW field given by
\begin{equation}
\Delta h^{TT}_{jk} = \frac{2}{R}\Delta\left(\ddot{\mathit{I}}^{TT}_{jk}\right) 
\end{equation} 
where $I_{ij}$ is the source mass-quadrupole moment, R is the distance to the source and $\Delta$ describes the net change in the quantity from early to late times. Gravitational waves with linear memory are important in case of unbound systems and have been studied studied in context of supernova explosion, asymmetric mass loss due to neutrino emission and gamma-ray-burst jets, \cite{Davies_Melvyn2002,Buonanno_Sigl2015,Kotake_Kei2009}
\par The linear memory produced by a system of N bodies with changing mass $M_A$ or velocities $V_A$ is given by the following general formula for by Thorn \cite{Thorne_Braginskii1987}

\begin{equation}\label{eq:1.2}
\Delta h^{TT}_{jk} = \Delta \sum_{A=1}^{N}\frac{4 M_A}{R\sqrt{1-v^2_A}}\left[\frac{v^j_A v^k_A}{1-\boldsymbol{v}_A\cdot \boldsymbol{N}}\right]
\end{equation}

where the masses are unbound in their initial and final states(or both). Here $\Delta$ refers to take difference between the final and initial values of the summations, and $\boldsymbol{N}$ is the unit vector that points from the source to the observer.  

@@

The other kind of memory is the nonlinear memory, which is also called Christodoulou memory. This was independently discovered by Christodoulou \cite{Christodoulou1991}, Payne \cite{Payne1894}, Blanchet and Damour \cite{Blanchet_Damour1992}. Christodoulou arises from a change in the radiative multiple moment that is sourced by the energy flux of the radiated gravitational wave. It can be understood in the following way. Consider the Einstein field equation in harmonic gauge. \cite{Thorne1980} 

@@

where $\frac{dE^{gw}}{dtd\Omega}$ is the GW energy flux, $n_{j}$ is a unit radial vector, the angle-bracket mean to average over several wavelengths, $h^{TT}_{ab}=\bar{h}^{TT}_{ab}$. When we apply standard Green's function to right hand side of equation \ref{eq:3.1a}, we obtain the following correction term to GW field \cite{Wiseman1991}

@@

where $T_{R}$ is the retarded time. This equation shows that part of the distant GW is sourced by loss of GW energy. The nonlinear memory occurs in any system that radiates GWs. These systems are usually considered to have a vanishing linear memory. The nonlinear memory [Eq.\ref{eq:1.6}] can be described by equation \ref{eq:1.2}, the formula for the linear memory if the unbound object in the system are taken to be the individual gravitons with energy $E_A = M_A/(1-v_A^2)^{1/2}$ and velocities $v_A^j = c\,{n'}_A^{j}$.\cite{Thorne1992} 
\par The Christodoulou memory is an interesting consequence of the nonlinearity of general relativity, it arises from the loss of GW energy from the system and the effect of this loss on the system's radiative mass multipole moments. The nonlinear memory effect arises from nonlinear interactions at 2.5 PN order but affects the gravitational waveform at leading (0PN) order. This memory effect causes a non oscillatory shift in the amplitude of $+$ polarization, which starts at early times and slowly grows during the inspiral. The PN waveform allow us to model the slow growth of memory during the inspiral phase of coalescence, but the most contribution to this nonlinear memory come from the merger phase where the PN model do not work. In this case the memory must be extracted from Numerical relativity simulations. However, it is quite challenging to extract the nonlinear memory accurately from numerical simulations of binary black holes. Numerical relativity simulations can most accurately compute the $(l, m) = (2, 2)$ mode of the waveform; but the nonlinear memory is present in only $m=0$ mode. These $m=0$ modes tend to be much smaller in magnitude and depends more sensitively on	initial conditions than other modes. For detailed discussion on this see section $\RomanNumeralCaps{5}$ in reference \cite{Favata2009}

@@

In terms of the $\psi_{lm}$ modes, the leading-order Christodoulou-memory modes are smaller than the $\psi_{22}$ modes by $five$ PN orders.The relative mode magnitudes(\cite{Favata2009}) are $\psi_{22} \approx 690 \psi_{20}\approx 4.5 \times 10^4 \psi_{40}$ at $x\approx1/5$. There are many other difficulties too in calculating memory from numerical relativity data. Construction of the $h_{lm}$ from $\psi_{lm}$ requires two integration constants to be specified. Choosing these integration constant to be zero \cite{Reisswig_Pollney2011} causes an artificial memory and a slope in the $h_{lm}$ modes. This artificial memory is unphysical and arises from the finite size of the initial separation and extraction radius.The initial burst of junk radiation also contribute to artificial memory.

@@

\par The corrections to the leading-oder formula for the nonlinear memory's contribution to 3PN order for non-spinning binaries has been computed by by Favata \cite{Favata2009}. However the effect of binary spins have not been explored these results.(\ashok{need to add why looking at spins is important})   
\par In this paper we calculate the nonlinear memory contribution to the + waveform polarization and look at the effect of spin contribution. In Sec. II, we describe the calculation required to compute memory contribution to the post-Newtonian waveform of quasi-circular, inspiralling binaries. In Sec. III, we describe the method used to extract memory contribution from NR simulations. Sec. IV we give the results which looks at the effect of spin on total nonlinear memory. We use PN expressions for memory for early time and for late time we use memory extracted from NR simulations.

@@

The mode decomposition of gravitational wave polarization is given by
\begin{equation}\label{eq:2.1}
	h_+ - \mathit{i}h_{\times} = \sum_{l=2}^{\infty}\sum_{m=-l}^{m=l}h^{lm} \,  _{-2}Y^{lm}(\Theta, \Phi)
\end{equation}
where $_{-2}Y^{lm}$ are spin-weighted spherical harmonics, the angles $(\Theta, \Phi)$ indicates the direction from the source to the observer. In a multipolar expansion of the GW field, the modes $h^{lm}$ are related to the radiative mass $U^{lm}$ and the current $V^{lm}$ multipole via
\begin{equation}\label{eq:2.2}
	h^{lm}= \frac{G}{\sqrt{2}\, R }\left[U^{lm}(T_R)-\frac{\mathit{i}}{c}V^{lm}(T_R)\right]
\end{equation}
Here $R$ is the distance from source to observer, $T_R$ is retarded time. We assume $c\,=\,G\,=\,1$ for all out calculations.
 The spin weighted spherical harmonics are defined in terms of the Wigner $d$ function by
\begin{equation}\label{eq:2.3}
	_{-2}Y^{lm}(\Theta, \Phi) = (-1)^2\sqrt{\frac{2l + 1}{4\pi}}d^{l}_{ms}(\Theta)e^{\mathit{i}m\Phi}
\end{equation} 
Here
\begin{align}\label{eq:2.4}
	d^{l}_{ms}&=\sqrt{(l+m)!(l-m)!(l+s)!(l-s)!}\\
	&\times \sum_{k=k_i}^{k_f}\frac{(-1)^k(\sin\frac{\Theta}{2})^{2k+s-1}(\cos\frac{\Theta}{2})^{1l+m-s-2k}}{k!(l+m-k)!(l-s-k)!(s-m+k)!}
\end{align}
where $k_i$ = max$(0, m-s)$ and $k_f$ = min$(l+m, l-s)$. \\
When the GW field is decomposed into spin-weighted spherical harmonic modes, the nonlinear memory can be shown to yield a correction to the radiative mass multipole moments that enters at 2.5PN and higher orders. For more detail about GW multipole expansion \cite{Thorne1980}. The correction to the radiative mass multipole moments is given by
\begin{equation}\label{eq:2.6}
	U^{(mem)}_{lm}=32\pi\frac{(l-2)!}{2(l+2)!}\int_{-\infty}^{T_R}dt\int d\Omega\frac{dE_{gw}}{dtd\Omega}(\Omega)Y^*_{lm}(\Omega)
\end{equation}
The radiative current moments $V_{lm}$ do not contribute to nonlinear memory. The GW energy flux can be computed from the GW stress-energy tensor and is given by \cite{Favata2010}
\begin{equation}\label{eq:2.7}
	\frac{dE_{gw}}{dtd\Omega} = R^2 T^{gw}_{00} = \frac{R^2}{32\pi}\langle\dot{h}^{TT}_{jk}\dot{h}^{TT}_{jk}\rangle =\frac{R^2}{16\pi}\langle\dot{h}^2_{+}+ \dot{h}^2_{\times}\rangle
\end{equation}
where the angled bracket mean to average over several wavelengths. The energy flux can be written in terms of $h_{lm}$ modes using equation \ref{eq:2.1} in the following way:
\begin{widetext}
\begin{equation}\label{eq:2.8}
		\frac{dE_{gw}}{dtd\Omega}=\frac{R^2}{16\pi}\sum_{l'=2}^{\infty}\sum_{l''=2}^{\infty}\sum_{m'=-l'}^{l'}\sum_{m''=-l''}^{l''}\langle\dot{h}_{l'm'}\dot{h}^*_{l''m''}\rangle _{-2}Y^{l'm'}(\theta,\phi)_{-2}Y^{l''m''*}(\theta, \phi)
\end{equation}
\end{widetext}

@@

	& \qquad\qquad\quad\left. + \frac{791\sqrt{1-4\eta}}{73728} + \frac{113265}{7340032}c_{\Theta}^4 + \eta\left(\frac{32497919}{16515072} - \frac{113\sqrt{1-4\eta}}{2048}\right)\right]

@@

	\includegraphics[width=3.5in]{../plots/PNmemorycontributionHpAlginedSpinAngular.pdf}

@@

	\includegraphics[width=3.5in]{../plots/PNmemorycontributionHpAntiAlginedSpinAngular.pdf}

@@

\section{Memory calculation using Numerical relativity data}
Numerical relativity simulations most commonly output the Newman-Penrose curvature component $\psi_4$.The two polarization states, $h_{+}$ and $h_{\times}$ of the gravitational wave are related to the curvature, expressed in terms 	of the complex Newman-Penrose scalar $\psi_4$ by 
	\psi_4 = \ddot{h}_+ - \mathit{i}\ddot{h}_{\times}
\end{equation}\
Given the Newman-Penrose scalar $\psi_4$ for a particular mode, we have to integrate twice in time to obtain $h_{+}$ and $h_{\times}$. It has long been noted that producing a strain, $h$, from the Newman-Penrose curvature component, $\psi_4$, typically results in a waveform with unphysical secular non-linear drift \cite{Berti2007}. The nonlinearity of drift indicates that this is not simply a result of two constants of integration involved in transformation. This nonlinearity is potentially caused by the fact that $\psi_4$ is typically extracted at a finite distance from the gravitating source. The strain $h$ is only related to $\psi_4$ at an infinite distance and also strictly valid in a particular gauge. As a result in numerical simulation, the finite distance calculation introduces a systematic error. In simulations \cite{Hannam2009, Reisswig2009}, which can possibly extract truly gauge-invariant waveform at future null infinity has not able able to get rid of the secular nonlinear drift.
\par It has been argued in ref. \cite{Reisswig_Pollney2011} that an important source of unphysical non-linear drift in numerical computation of gravitational wave strain lies in the transformation of the measured data to the strain $h$  which generally involves an integration in time. The output of the numerical simulation is a discretely sampled time series of finite duration, incorporating some component of unresolved frequencies due to numerical error. This can lead to an uncontrollable non-linear drift if the integration is performed in the time domain. The method described in ref \cite{Reisswig_Pollney2011} gives us a way to get rid of this unphysical behavior by eliminating the lower frequencies.    
\par The memory contribution to $h_{+}$ is given by \cite{Favata2010}

@@

\subsection{Memory calculation from SXS Rumerical Relativity data}
The publicly available SXS catalog provides data for Newman-Penrose scalar component $\psi_4$, decomposed into spin weighted spherical harmonics, extrapolated to infinite radius. We used this data and  performed a frequency domain time integration as described in equation \ref{3.6}, while removing the unphysical frequency by choosing a lower cutoff as show in section \ref{FFI}. The cutoff frequency is chosen as $\omega_0/10$ where $\omega_0$ is the instantaneous GW angular frequency. We first begin by separating the complex Newman-Perose scalar $\psi_4$ data into phase and amplitude part. We compute instantaneous GW angular frequency from this by taking the first time derivative of phase corresponding to $\psi_4$. The instantaneous GW frequency often looks noisy at the beginning of the simulations due to numerical issues. We remove this noise data by choosing a lower and upper cutoff time. Once we find the upper and lower cutoff time, we can apply a frequency domain time integration for $\psi_4$ by choosing a cutoff frequency. After filtering the noise creating frequency we can use equation \ref{eq:3.2} to compute the memory contribution to "plus" polarization of GW signal. The SXS catalog gives data for different parameter values like mass ratio, dimensionless spin parameter etc. The Christodoulou memory computed from Numerical Relativity data starts from zeros using equation \ref{eq:3.2} due to finite simulation time. Chrostodoulou memory is hereditary, hence it's entire past should be included. This can be done by combining the PN expression during early inspiral and numerical relativity results for merger part. The method used to connect the memory accumulated during entire life time of BBH coalesce is described in the next section. 
\subsection{Compleate Christodoulou memory during BBH coalescence} 
The PN waveform are given as a function of parameter $x$ which is related to time by equation \ref{eq:2.14}. This equation can be integrated to obtain a relation between time $t$ and parameter $x$. The Christodoulou memory reaches it's peak value at the merger. We choose the values of integration constants such that the peak in $h_+^{(mem)}$ is reached at time $t=0$. For comparison we show the time derivative of $h_+^{(mem)}$ from Numerically obtained results and the analytical expressions in figure \ref{fig:derivative}. The slope as seen from figure \ref{fig:derivative} rises asymptotically. We use our PN results for computing the memory during early inspiral and attach them to NR results. We compare this memory for different mass ratio and different values of dimensionless spin parameter. Figure \ref{fig:differentmassratio} show the Christodoulou memory for different mass ratios, which shows that the Christodoulou tends to decrease with higher mass ration. The effect of dimensionless spin is shown in figure \ref{fig:differntchivalues}. It can be seen that the spin magnitude has significant effect on nonlinear memory, higher spin magnitude, in case of spin aligned with orbital angular momentum tends to increase the memory contribution. When the spins are anti aligned with orbital angular momentum, the memory tend to decrease the Christodoulou memory. These results show it important to consider the spin contribution when studying the Christodoulou memory in gravitational wave signals.

>>>>

Gravitational waves, when they pass through a pair of freeling falling test masses cause an oscillatory motion. This oscillatory motion after the gravitational wave has passed through it, does not bring back the pair of test particle to their original location. This permanent displacement is called gravitational wave memory effect. This displacement is caused by a permanent deformation of spacetime geometry caused by passing gravitational waves. The memory effect associated with gravitational waves is accumulated effect over the entire lifehistory of the source. The permanent deformation that it leaves in spacetime geometry can be described in two ways: The Linear memory, which is caused by net change in the time derivative of source of one more multipole moment and the Nonlinear memory, which is caused by radiatiated gravitational waves acting as source for generating secondary GWs. The linear memory is important for unbound astrophysical system and have been studied in context of supernova explosion, asymmetric mass loss due to neutrino emission, gamma-ray-burst jets etc \cite{Davies_Melvyn2002,Buonanno_Sigl2015,Kotake_Kei2009}. At leading order in a Post-Newtonian (PN) expansion, the linear memory causes a net change in the GW field given by
\begin{equation}
\Delta h^{TT}_{jk} = \frac{2}{R}\Delta\left(\ddot{\mathit{I}}^{TT}_{jk}\right) 
\end{equation} 
where $I_{ij}$ is the source mass-quadrupole moment, R is the distance to the source and $\Delta$ describes the net change in the quantity from early to late times. 
\par A general formula for the linear memory produced by an unbound system of N particle as a result of  change in  it’s mass $M_A$ and velocity configuration $v_A$ is given by Thorn \cite{Thorne_Braginskii1987} as shown below.
\begin{equation}\label{eq:1.2}
\Delta h^{TT}_{jk} = \Delta \sum_{A=1}^{N}\frac{4 M_A}{R\sqrt{1-v^2_A}}\left[\frac{v^j_A v^k_A}{1-\boldsymbol{v}_A\cdot \boldsymbol{N}}\right]
\end{equation}

 Here $\Delta$ refers to take difference between the final and initial values of the summations, and $\boldsymbol{N}$ is the unit vector that points from the source to the observer.  

@@

The other kind of memory is the nonlinear memory, which is also called Christodoulou memory. It arises as results of change in the radiative multipole moment that is sourced by the energy flux of the radiated GWs. It can be understood by looking at Eisntein's equations in harmonic gauge and considering the leading order nonlinear terms. The equation below shows a wave equation for spacetime metric in harmonic gauge with different source terms appearing on the right. \cite{Thorne1980} 

@@

where $\frac{dE^{gw}}{dtd\Omega}$ is the GW energy flux, $n_{j}$ is a unit radial vector, the angle-bracket mean to average over several wavelengths, $h^{TT}_{ab}=\bar{h}^{TT}_{ab}$. The standard procedure to solve a wave equation with source is by using Green’s function. When applied to  Eq. \ref{eq:3.1a} obtains a correction term to gravitational field sourced by the Landau-Lifshitz term is shown below \cite{Wiseman1991}

@@

where $T_{R}$ is the retarded time. This equation shows that part of the distant GW is sourced by loss of GW energy. This was independently discovered by Christodoulou \cite{Christodoulou1991}, Payne \cite{Payne1894}, Blanchet and Damour \cite{Blanchet_Damour1992}. The nonlinear memory occurs in any system that radiates GWs. These systems are usually considered to have a vanishing linear memory. The nonlinear memory [Eq.\ref{eq:1.6}] can be described by equation \ref{eq:1.2}, the formula for the linear memory if the unbound object in the system are taken to be the individual gravitons with energy $E_A = M_A/(1-v_A^2)^{1/2}$ and velocities $v_A^j = c\,{n'}_A^{j}$.\cite{Thorne1992} 
\par The Christodoulou memory effect arises from nonlinear interactions at 2.5 PN order but affects the gravitational waveform at leading (0PN) order. The effect of Christodoulou memory on GW waveform is a non oscillatory shift in the amplitude of $+$ polarization, which starts at early times and slowly grows during the inspiral. The PN waveform allow us to model the slow growth of memory during the inspiral phase of coalescence, but the most contribution to this nonlinear memory come from the merger phase where the PN model do not work. In this case the memory must be extracted from Numerical relativity simulations. However, it is quite challenging to extract the nonlinear memory accurately from numerical simulations of binary black holes. Numerical relativity simulations can most accurately compute the $(l, m) = (2, 2)$ mode of the waveform; but the nonlinear memory is present in only $m=0$ mode. These $m=0$ modes are much smaller in magnitude and depends more sensitively on	initial conditions than other modes. For detailed discussion on this see section $\RomanNumeralCaps{5}$ in reference \cite{Favata2009}

@@

In terms of the $\psi_{lm}$ modes, the leading-order Christodoulou-memory modes are smaller than the $\psi_{22}$ modes by $five$ PN orders. The relative mode magnitudes(\cite{Favata2009}) are $\psi_{22} \approx 690 \psi_{20}\approx 4.5 \times 10^4 \psi_{40}$ at $x\approx1/5$. There are other difficulties too in calculating memory from numerical relativity data. Construction of the $h_{lm}$ from $\psi_{lm}$ requires two integration constants to be specified. Choosing these integration constant to be zero \cite{Reisswig_Pollney2011}  does not make the signal oscillate around zero. So there is an artificial memory in the $h_{lm}$ modes which is caused by numerical noise. This artificial memory is unphysical and arises from the finite size of the initial separation and extraction radius. The initial burst of junk radiation also contribute to artificial memory.

@@

\par The corrections to the leading-oder formula for the nonlinear memory's contribution to 3PN order for non-spinning binaries has been computed by by Favata \cite{Favata2009}. However the effect of binary spins have not been explored in these results. Astrophysical black holes are formed by gravitation collapse of massive star when when they run out of their nuclear fuel in the core. These massive stars are likely to posses spin angular momentum, as a result the black hole formed would also be spinning. The black holes can also accrete matter from surrounding and as a result spin up during the process. It is highly likely that there is a large population of spinning binaries in the universe.      
\par In this paper we calculate the nonlinear memory contribution to the + waveform polarization and look at the effect of spin contribution. In Sec. II, we describe the calculation required to compute memory contribution to the post-Newtonian waveform of quasi-circular, inspiralling binaries. In Sec. III, we describe the method used to extract memory contribution from NR simulations. Sec. IV we give the results which looks at the effect of spin on total nonlinear memory. We use PN expressions for memory for early time and for late time we use memory extracted from NR simulations.

@@

The mode decomposition in terms spin-weighted spherical harmonics for gravitational wave polarization is given by
\begin{equation}\label{eq:2.1}
	h_+ - \mathit{i}h_{\times} = \sum_{l=2}^{\infty}\sum_{m=-l}^{m=l}h^{lm} \,  _{-2}Y^{lm}(\Theta, \Phi)
\end{equation}
where $_{-2}Y^{lm}$ are spin-weighted spherical harmonics, the angles $(\Theta, \Phi)$ indicates the direction from the source to the observer. In a multipolar expansion of the GW field, the modes $h^{lm}$ are related to the radiative mass $U^{lm}$ and the current $V^{lm}$ multipole via
\begin{equation}\label{eq:2.2}
	h^{lm}= \frac{G}{\sqrt{2}\, R }\left[U^{lm}(T_R)-\frac{\mathit{i}}{c}V^{lm}(T_R)\right]
\end{equation}
Here $R$ is the distance from source to observer, $T_R$ is retarded time. We assume $c\,=\,G\,=\,1$ for all out calculations.
 The spin weighted spherical harmonics are defined in terms of the Wigner $d$ function by
\begin{equation}\label{eq:2.3}
	_{-2}Y^{lm}(\Theta, \Phi) = (-1)^2\sqrt{\frac{2l + 1}{4\pi}}d^{l}_{ms}(\Theta)e^{\mathit{i}m\Phi}
\end{equation} 
Here
\begin{align}\label{eq:2.4}
	d^{l}_{ms}&=\sqrt{(l+m)!(l-m)!(l+s)!(l-s)!}\\
	&\times \sum_{k=k_i}^{k_f}\frac{(-1)^k(\sin\frac{\Theta}{2})^{2k+s-1}(\cos\frac{\Theta}{2})^{1l+m-s-2k}}{k!(l+m-k)!(l-s-k)!(s-m+k)!}
\end{align}
where $k_i$ = max$(0, m-s)$ and $k_f$ = min$(l+m, l-s)$. \\
When the GW field is decomposed into spin-weighted spherical harmonic modes, the nonlinear memory can be shown to yield a correction to the radiative mass multipole moments that enters at 2.5PN and higher orders. For more detail about GW multipole expansion \cite{Thorne1980}. The correction to the radiative mass multipole moments is given by
\begin{equation}\label{eq:2.6}
	U^{(mem)}_{lm}=32\pi\frac{(l-2)!}{2(l+2)!}\int_{-\infty}^{T_R}dt\int d\Omega\frac{dE_{gw}}{dtd\Omega}(\Omega)Y^*_{lm}(\Omega)
\end{equation}
The radiative current moments $V_{lm}$ do not contribute to nonlinear memory. The GW energy flux can be computed from the GW stress-energy tensor and is given by \cite{Favata2010}
\begin{equation}\label{eq:2.7}
	\frac{dE_{gw}}{dtd\Omega} = R^2 T^{gw}_{00} = \frac{R^2}{32\pi}\langle\dot{h}^{TT}_{jk}\dot{h}^{TT}_{jk}\rangle =\frac{R^2}{16\pi}\langle\dot{h}^2_{+}+ \dot{h}^2_{\times}\rangle
\end{equation}
where the angled bracket mean to average over several wavelengths. The energy flux can be written in terms of $h_{lm}$ modes using equation \ref{eq:2.1} in the following way:
\begin{widetext}
\begin{equation}\label{eq:2.8}
		\frac{dE_{gw}}{dtd\Omega}=\frac{R^2}{16\pi}\sum_{l'=2}^{\infty}\sum_{l''=2}^{\infty}\sum_{m'=-l'}^{l'}\sum_{m''=-l''}^{l''}\langle\dot{h}_{l'm'}\dot{h}^*_{l''m''}\rangle _{-2}Y^{l'm'}(\theta,\phi)_{-2}Y^{l''m''*}(\theta, \phi)
\end{equation}
\end{widetext}

@@

	& \qquad\qquad\quad\left. + \frac{791\sqrt{1-4\eta}}{73728} + \frac{113265}{7340032}c_{\Theta}^4 + \eta\left(\frac{32497919}{16515072} - \frac{113\sqrt{1-4\eta}}{2048} -\frac{8200699}{8257536}c_{\Theta}^2 - \frac{791\sqrt{1-4\eta}}{18432}c_{\Theta}^2 - \frac{701315}{5505024}c_{\Theta}^4\right)\nonumber\right.\\
	& \qquad\qquad\quad\left. + \eta^2\left(-\frac{142301}{36864} + \frac{40501}{18432}c_{\Theta}^2 + \frac{1075}{4096}c_{\Theta}^4\right)\right]\chiaSqr + 
	\left[\frac{4675\pi}{1152}-\frac{9\pi\sqrt{1-4\pi}}{2048} + \frac{275\pi}{1152}c_{\Theta}^2 - \frac{7\pi\sqrt{1-4\pi}}{2048}c_{\Theta}^2 + \eta\left(-\frac{947\pi}{288}\right.\right.\nonumber\\
	& \qquad\qquad\quad + \left.\left.\frac{9\pi\sqrt{1-4\eta}}{512} -\frac{91\pi}{288}c_{\Theta}^2 + \frac{7\pi\sqrt{1-4\eta}}{512}c_{\Theta}^2\right)\right]\chisdL + \left[-\frac{14443855}{66060288} + \frac{113\sqrt{1-4\eta}}{8192} + \frac{3716075}{33030144}c_{\Theta}^2 + \frac{791\sqrt{1-4\eta}}{73728}c_{\Theta}^2\right.\nonumber\\
	& \qquad\qquad\quad + \frac{113265}{7340032}c_{\Theta}^4 + \eta^3\left(-\frac{11}{192} - \frac{11}{384}c_{\Theta}^2\right) + \eta\left(\frac{15798397}{2752512} - \frac{33\sqrt{1-4\eta}}{512} + \frac{921469}{4128768}c_{\Theta}^2 - \frac{77\sqrt{1-4\eta}}{1536} - \frac{184955}{2752512}c_{\Theta}^4\right)\nonumber\\
	& \qquad\qquad\quad \left.+ \eta^2\left(-\frac{9507487}{4128768} + \frac{19\sqrt{1-4\eta}}{512} - \frac{676133}{2064384}c_{\Theta}^2 +\frac{133\sqrt{1-4\eta}}{4608}c_{\Theta}^2+ \frac{11745}{458752}c_{\Theta}^4\right)\right]\chisSqr + \left[-\frac{9\pi}{2048} + \frac{4675\pi\sqrt{1-4\eta}}{1152}\right.\nonumber\\
	& \qquad\qquad\quad - \frac{7\pi}{2048}c_{\Theta}^2 + \frac{275\pi\sqrt{1-4\eta}}{1152}c_{\Theta}^2 + \eta\left(\frac{9\pi}{512} + \frac{7\pi}{512}c_{\Theta}^2\right) + \left(\frac{113}{4096} -\frac{14443855\sqrt{1-4\eta}}{33030144} + \frac{791}{36864}c_{\Theta}^2 + \frac{3716075\sqrt{1-4\eta}}{16515072}c_{\Theta}^2\right.\nonumber\\
	 & \qquad\qquad\quad +\frac{113265\sqrt{1-4\eta}}{3670016}c_{\Theta}^4 + \eta^2\left(\frac{33}{128} + \frac{11\sqrt{1-4\eta}}{384} + \frac{77}{384}c_{\Theta}^2 + \frac{11\sqrt{1-4\eta}}{768}c_{\Theta}^2\right) + \eta\left(-\frac{179}{1024} + \frac{56422223\sqrt{1-4\eta}}{8257536}\right.\nonumber\\
	 & \qquad\qquad\quad \left.\left.\left.\left. -\frac{1253}{9216}c_{\Theta}^2 -\frac{440281\sqrt{1-4\eta}}{1376256}c_{\Theta}^2 -\frac{17415\sqrt{1-4\eta}}{131072}c_{\Theta}^4\right)\right)\chisdL\right]\right\}

@@

	\includegraphics[width=3.5in]{../plots/PNmemoryFavata.pdf}

@@

@@

\begin{figure}
	\includegraphics[width=3.5in]{../plots/PNmemorycontributionHpAlginedSpinAngular.pdf}
	\includegraphics[width=3.5in]{../plots/PNmemorycontributionHpAntiAlginedSpinAngular.pdf}
	\caption{Dependence of post-Newtonian (PN) corrections to the Christodoulou memory on the binary inclination and orbital separation. The plots above show the memory contribution to $H_+$ for different dimensionless spin parameter $\chi$ values (the polar angle to the observer at 𝛩 = 0 points along the binary orbital angular momentum) for $x = 1/5$ and $\eta=1/4 $ . The plot to the right shows $x H_+$ as a function of parameter x. Increasing the magnitude of 𝜒 does not change the angular dependence, but does tend to increase the magnitude of the memory. 
	}
\end{figure}
\section{Memory calculation using Numerical relativity data}
Numerical relativity simulations most commonly output the Newman-Penrose curvature component $\psi_4$.The two polarization states, $h_{+}$ and $h_{\times}$ of the gravitational wave are related to the curvature, expressed in terms 	of the complex Newman-Penrose scalar $\psi_4$ by 
\begin{equation}
	\psi_4 = \ddot{h}_+ - \mathit{i}\ddot{h}_{\times}
\end{equation}\
Given the Newman-Penrose scalar $\psi_4$ for a particular mode, we have to integrate twice in time to obtain $h_{+}$ and $h_{\times}$. It has long been known that producing a strain, $h$, from the Newman-Penrose curvature component, $\psi_4$, typically results in a waveform with unphysical secular non-linear drift \cite{Berti2007}. The nonlinearity of drift indicates that this is not simply a result of two constants of integration involved in transformation. This nonlinearity is potentially caused by the fact that $\psi_4$ is typically extracted at a finite distance from the gravitating source. The strain $h$ is only related to $\psi_4$ at an infinite distance and also strictly valid in a particular gauge. As a result in numerical simulation, the finite distance calculation introduces a systematic error. In simulations \cite{Hannam2009, Reisswig2009}, which can possibly extract truly gauge-invariant waveform at future null infinity has not able able to get rid of the secular nonlinear drift.
\par It has been argued in ref. \cite{Reisswig_Pollney2011} that an important source of unphysical non-linear drift in numerical computation of gravitational wave strain lies in the transformation of the measured data to the strain $h$  which generally involves an integration in time. The output of the numerical simulation is a discretely sampled time series of finite duration, incorporating some component of unresolved frequencies due to numerical error. This can lead to an uncontrollable non-linear drift if the integration is performed in the time domain. The method described in ref \cite{Reisswig_Pollney2011} gives us a way to get rid of this unphysical behavior by eliminating the lower frequencies.    
\par The memory contribution to $h_{+}$ is given by \cite{Favata2010}

@@

\subsection{Memory calculation from SXS Rumerical Relativity data}
The publicly available SXS catalog provides data for Newman-Penrose scalar component $\psi_4$, decomposed into spin weighted spherical harmonics, extrapolated to infinite radius. We used this data and  performed a frequency domain time integration as described in equation \ref{3.6}, while removing the unphysical frequency by choosing a lower cutoff as show in section \ref{FFI}. The cutoff frequency is chosen as $\omega_0/10$ where $\omega_0$ is the minimum instantaneous GW angular frequency. We first begin by separating the complex Newman-Perose scalar $\psi_4$ data into phase and amplitude part. We compute instantaneous GW angular frequency from this by taking the first time derivative of phase corresponding to $\psi_4$. The instantaneous GW frequency often looks noisy at the beginning of the simulations due to numerical issues. We remove this noise data by choosing a lower and upper cutoff time. Once we find the upper and lower cutoff time, we can apply a frequency domain time integration for $\psi_4$ by choosing a cutoff frequency. After filtering the noise creating frequency we can use equation \ref{eq:3.2} to compute the memory contribution to "plus" polarization of GW signal. The SXS catalog gives data for different parameter values like mass ratio, dimensionless spin parameter etc. The Christodoulou memory computed from Numerical Relativity data starts from zeros using equation \ref{eq:3.2} due to finite simulation time. Chrostodoulou memory is hereditary, hence it's entire past should be included. This can be done by combining the PN expression during early inspiral and numerical relativity results for merger part. The method used to connect the memory accumulated during entire life time of BBH coalesce is described in the next section. 
\subsection{Compleate Christodoulou memory during BBH coalescence} 
The PN waveform for Christodoulou are given as a function of parameter $x$ which is related to time by equation \ref{eq:2.14}. This equation can be integrated to obtain a relation between time $t$ and parameter $x$. The Christodoulou memory reaches it's peak value at the merger. We choose the values of integration constants such that the peak in $h_+^{(mem)}$ is reached at time $t=0$. We use our PN results for computing the memory during early inspiral and attach them to NR results near the late inspiral. We compare this memory for different mass ratio and different values of dimensionless spin parameter. Figure \ref{fig:differentmassratio} show the Christodoulou memory for different mass ratios, which shows that the Christodoulou tends to decrease with higher mass ration. The effect of dimensionless spin is shown in figure \ref{fig:differntchivalues}. It can be seen that the spin magnitude has significant effect on nonlinear memory, higher spin magnitude, in case of spin aligned with orbital angular momentum tends to increase the memory contribution. When the spins are anti aligned with orbital angular momentum, it tend to decrease the Christodoulou memory. These results show that it is important to consider the spin contribution when studying the Christodoulou memory in gravitational wave signals. 
\subsection{CONCLUSIONS}
We find that spin effects are significantly important for observations that aims at detecting Chrisodoulou memory.

\section{Acknowledgments}
We thank the West Virginia Univerisity. This work was support by ...
